### 2016-04-13 tabular.tex
Almost finished the ecological introduction

 \begin{tabular}{ | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | }
  \hline
  \multicolumn{7}{|c|}{Faszpina} \\
  \hline
 	& A & B & C & D & E & F\parbox{0pt}{\rule{0pt}{2ex+\baselineskip}}\\
 	\hline
 	A & 0  & 0 & 0 & 0 & $\frac{1}{3}$ & 0\parbox{0pt}{\rule{0pt}{2ex+\baselineskip}}\\
 	\hline

 	\hline
 \end{tabular}

  \begin{tabular}{ | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | >{\centering\arraybackslash}m{5pt} | }
   \hline
   \multicolumn{7}{|c|}{Faszpina} \\
   \hline
  	& A & B & C & D & E & F\parbox{0pt}{\rule{0pt}{2ex+\baselineskip}}\\
  	\hline
  	A & 0  & 0 & 0 & 0 & $\frac{1}{3}$ & 0\parbox{0pt}{\rule{0pt}{2ex+\baselineskip}}\\
  	\hline
  	B & 0 & 0 & 0 & 0 & $\frac{1}{3}$ & $\frac{1}{3}$\parbox{0pt}{\rule{0pt}{2ex+\baselineskip}}\\
  	\hline
  	C & 0 & 0 & 0 & 0 & $\frac{1}{3}$ & $\frac{1}{3}$\parbox{0pt}{\rule{0pt}{2ex+\baselineskip}}\\
  	\hline
  	D & 0 & 0 & 0 & 0 & 0 & $\frac{1}{3}$\\[6pt] 
  	\hline
  	E & 1 & $\frac{1}{2}$ & $\frac{1}{2}$ & 0 & 0 & 0\parbox{0pt}{\rule{0pt}{2ex+\baselineskip}}\\
  	\hline
  	F & 0 & $\frac{1}{2}$ & $\frac{1}{2}$ & 1 & 0 & 0\parbox{0pt}{\rule{0pt}{2ex+\baselineskip}}\\
  	\hline

  \end{tabular}

  % sima mátrix

  \[ 
  \begin{bmatrix}
  0 & 0 & 0 & 0 & \frac{1}{3} & 0 \\[6pt]
  0 & 0 & 0 & 0 & \frac{1}{3} & \frac{1}{3}\\[6pt] 
  0 & 0 & 0 & 0 & \frac{1}{3} & \frac{1}{3}\\[6pt] 
  0 & 0 & 0 & 0 & 0 & \frac{1}{3}\\[6pt] 
  1 & \frac{1}{2} & \frac{1}{2} & 0 & 0 & 0\\[6pt] 
  0 & \frac{1}{2} & \frac{1}{2} & 1 & 0 & 0
  \end{bmatrix}
  \quad
  \begin{bmatrix}
  \frac{1}{3} & \frac{1}{6} & \frac{1}{6} & 0 & 0 & 0 \\[6pt] 
  \frac{1}{3} & \frac{1}{3} & \frac{1}{3} & \frac{1}{3} & 0 & 0\\[6pt] 
  \frac{1}{3} & \frac{1}{3} & \frac{1}{3} & \frac{1}{3} & 0 & 0\\[6pt] 
  0 & \frac{1}{6} & \frac{1}{6} & \frac{1}{3} & 0 & 0\\[6pt] 
  0 & 0 & 0 & 0 & \frac{2}{3} & \frac{1}{3}\\[6pt] 
  0 & 0 & 0 & 0 & \frac{1}{3} & \frac{2}{3}
  \end{bmatrix}
  \]